GSEA datasets for FDR benchmarking

### 2017-12-08 datasets/GSEA/gsea.Rmd
Start adding mouse datasets

library(EnsDb.Hsapiens.v86)
gnType <-  genes(EnsDb.Hsapiens.v86, column="gene_biotype")
dsdObject <- dsdObject[rownames(dsdObject) %in% gnType[gnType$gene_biotype == "protein_coding"]$gene_id,]
dsdObject <- DESeq(dsdObject)
res <- results( dsdObject, lfcThreshold=1 )

@@

```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}

library(ggplot2)
library(magrittr)
library(cowplot)
library(R.utils)
library(scales)
sourceDirectory("../R/")

@@

```{r runningBenckmark, message=FALSE}
designBen <- initializeBenchDesign()
res <- dplyr:::select( goRes, c("over_represented_pvalue", "numInCat") ) %>%
  dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
res$pval <- pmin( res$pval, 1 )
head(res)
sGSEA <- designBen %>% buildBench( res, ftCols="ind_covariate")
elementMetadata( rowData(sGSEA) )
assayNames(sGSEA) <- "qvalue"
sGSEA <- addDefaultMetrics( sGSEA )
```

@@

stopifnot( all(assays(sGSEA)[["qvalue"]][,"unadjusted"] == pmin(goRes$over_represented_pvalue, 1) ) )

sGSEA <- estimatePerformanceMetrics(sGSEA, addColData=TRUE)
rejection_scatter_bins( sGSEA, "ind_covariate", threshold=15, bins=4, as_fraction=TRUE )

@@

>>>>

library(EnsDb.Hsapiens.v86)
gnType <-  genes(EnsDb.Hsapiens.v86, column="gene_biotype")

dsdObject <- dsdObject[rownames(dsdObject) %in% gnType[gnType$gene_biotype == "protein_coding"]$gene_id,]
dsdObject <- DESeq(dsdObject)
res <- results( dsdObject, lfcThreshold=1 )

@@

```{r informativeCovariate, message=FALSE, fig.height=3.5, message=FALSE, fig.width=4.5}
library(ggplot2)
library(magrittr)
library(cowplot)
library(R.utils)
library(scales)
sourceDirectory("../R/")

@@

```{r runningBenckmark, message=FALSE}
designBen <- initializeBenchDesign()
res <- dplyr:::select( goRes, c("over_represented_pvalue", "numInCat") ) %>%
  dplyr:::rename( pval = over_represented_pvalue, ind_covariate = numInCat )
res$pval <- pmin( res$pval, 1 )
sGSEA <- designBen %>% buildBench( res )
assayNames(sGSEA) <- "qvalue"
sGSEA <- addDefaultMetrics( sGSEA )
```

@@

stopifnot( all(assays(sGSEA)[["qvalue"]][,"unadjusted"] == pmin(goRes$over_represented_pvalue, 1) ) )
rowData(sGSEA)$numInCat <- goRes$numInCat

rejection_scatter_bins( sGSEA, "numInCat", threshold=15, bins=4, as_fraction=TRUE )
```

Plotting the methods overlap

```{r}

sGSEASub <- sGSEA[,!apply( is.na( assays(sGSEA)[["qvalue"]] ), 2, all )]
sGSEASub <- sGSEASub[,-which(grepl("ihw", colData( sGSEASub )$blabel) & colData( sGSEASub )$param.alpha != 0.1)]
sGSEASub <- sGSEASub[,-which( grepl("bl", colData( sGSEASub)$blabel) & 
                   colData(sGSEASub)$param.smooth.df != 2L )]
sGSEASub <- sGSEASub[,!colData(sGSEASub)$blabel %in% "unadjusted"]

plotMethodsOverlap(sGSEASub, order.by = "freq", nsets=10)

@@

# Same analysis but using the HSC-MPP dataset from Nina

```{r}
dsdMouse <- "dsdHSCMPP.RData"
if( !file.exists( dsdMouse ) ){
  download.file("https://www.dropbox.com/s/spo1pifum5k72rn/dseHSCMPP.RData?dl=1", destfile="dsdHSCMPP.RData")
}
load(dsdMouse)
dseHSCMPP <- DESeq( dseHSCMPP)

res <- results( dseHSCMPP, contrast=c("conditions", "HSC", "MPP1") )
genes <- as.numeric( res$padj < 0.1 )
names( genes ) <- rownames( res )
genes[is.na( genes )] <- 0

### preparing gene length info and goSets ###
library("EnsDb.Mmusculus.v75")

pwf <- nullp(genes,  )
goRes <- goseq( pwf, gene2cat=goSets )
